\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-10-16 fully-abstract-comp/popl2019/paper.tex

\begin{figure}[htb]
  \centering
{\footnotesize
\[
  \begin{array}{r >{$}p{3.5cm}<{$} l >{$}p{3.5cm}<{$} l l}
    1: & \tmove{\rtmp{1}}{42} & 10: &\tcseal{\rretd}{\rtmp{1}} & 19: &\tcca{\rtmp{2}}{5} \\
    2: &\tstore{\rstk}{\rtmp{1}} & 11: &\tmove{\rretc}{\pcreg} & 20: &\tjnz{\rtmp{2}}{\rtmp{1}}\\
    3: &\tcca{\rstk}{(-1)} & 12: & \tcca{\rretc}{5} & 21: &\tcca{\rtmp{2}}{1} \\
    4: &\tgeta{\rtmp{1}}{\rstk} & 13: & \tcseal{\rretc}{\rtmp{1}} & 22: &   \tjmp{\rtmp{2}} \\
    5: &\tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}} & 14: & \tmove{\rtmp{1}}{0} & 23: &    \tfail \\
    6: &\tmove{\rtmp{1}}{\pcreg} & 15: & \txjmp{r_1}{r_2} & 24: &   \tsplice{\rstk}{\rstk}{\rdata} \\
    7: &\tcca{\rtmp{1}}{(\offpc - 5)} & 16: & \tgetb{\rtmp{1}}{\rstk} & 25: &    \tcca{\rstk}{1}\\
    8: &\tload{\rtmp{1}}{\rtmp{1}} & 17: &\tminus{\rtmp{1}}{\rtmp{1}}{\stkb} & 26: &    \tmove{\rtmp{2}}{0} \\
    9: &\tcca{\rtmp{1}}{\offsigma} & 18: &\tmove{\rtmp{2}}{\pcreg} & &
  \end{array}
\]
}
\caption{
  The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
  Variable $\offpc$ is the offset from the first instruction of the call to the set of seals, and $\offsigma$ is the offset to the seal within the set.
  Instructions 1-3: ensure that the stack is non-empty.
  Instructions 4-5: split stack pointer in two (stack pointer for callee done).
  Instructions 6-9: Load the seal designated for this call.
  Instruction 10: Seal capability for private stack to create data return pointer.
  Instructions 11-13: Seal capability for return point to create code return pointer.
  Instructions 14-15: Clean up temporary register and jump.
  Instructions 16-26: The return code.
  Instructions 16-23: check the stack base.
  Instruction 24: join the stack pointer.
  Instructions 25-26: pop 42 from the stack.
%The variable $\offpc$ is the offset from the first instruction of the call to the set of seals it uses. The variable $\offsigma$ is the offset in the set of seals. The constant $\stkb$ is the globally agreed on stack base. There are a number of ``magic'' numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack. Line 7: $-5$, offset to the first instruction (from the point the capability was moved out of the $\pcreg$-register). Line 12: $5$, offset to the return address. Line 19: $5$, offset to fail. Line 21: offset to address after fail.
}
  \label{fig:call-code}
\end{figure}

>>>>

\begin{figure}[htb]
  \centering
{\footnotesize
\[
  \begin{array}{r >{$}p{5.1cm}<{$} r l}
\multicolumn{2}{l}{\text{// Ensure non-empty stack.}}            & \multicolumn{2}{l}{\text{// Clear tmp registers and jump.}}\\
1 :& \tmove{\rtmp{1}}{42}                                        &14: & \tmove{\rtmp{1}}{0}\\
2 :& \tstore{\rstk}{\rtmp{1}}                                    &15: & \txjmp{r_1}{r_2}\\
3 :& \tcca{\rstk}{(-1)}                                          &\multicolumn{2}{l}{\text{// The following is the return code.}}\\
\multicolumn{2}{l}{\text{// Split stack in local stack frame and unused.}}      &\multicolumn{2}{l}{\text{// Check that returned stack pointer has base $\stkb$.}}\\
4 :& \tgeta{\rtmp{1}}{\rstk}                                     &16: & \tgetb{\rtmp{1}}{\rstk}\\
5 :& \tsplit{\rstk}{\rretd}{\rstk}{\rtmp{1}}                     &17: & \tminus{\rtmp{1}}{\rtmp{1}}{\stkb}\\
\multicolumn{2}{l}{\text{// Load the call seal.}}                &18: & \tmove{\rtmp{2}}{\pcreg}\\
6 :& \tmove{\rtmp{1}}{\pcreg}                                    &19: & \tcca{\rtmp{2}}{5}\\
7 :& \tcca{\rtmp{1}}{(\offpc - 5)}                               &20: & \tjnz{\rtmp{2}}{\rtmp{1}}\\
8 :& \tload{\rtmp{1}}{\rtmp{1}}                                  &21: & \rtmp{2}{1}\\
9 :& \tcca{\rtmp{1}}{\offsigma}                                  &22: & \tjmp{\rtmp{2}}\\
\multicolumn{2}{l}{\text{// Seal the local stack frame.}}        &23: & \tfail \\
10:& \tcseal{\rretd}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Splice with capability for local stack frame.}}\\
\multicolumn{2}{l}{\text{// Construct code return pointer.}}     &24: & \tsplice{\rstk}{\rstk}{\rdata} \\
11:& \tmove{\rretc}{\pcreg}                                      &\multicolumn{2}{l}{\text{// Pop 42 from the stack}}\\  
12:& \tcca{\rretc}{5}                                            &25: & \tcca{\rstk}{1}\\
13:& \tcseal{\rretc}{\rtmp{1}}                                   &\multicolumn{2}{l}{\text{// Clear tmp register}}\\
      &    &26: & \tmove{\rtmp{2}}{0}\\
  \end{array}
\]
}
\caption{
  % The instructions that constitutes a call. Register $r_1$ and $r_2$ are the registers where the sealed capability pair to the callee resides.
  % Variable $\offpc$ is the offset from the first instruction of the call to the set of seals, and $\offsigma$ is the offset to the seal within the set.
  % Instructions 1-3: ensure that the stack is non-empty.
  % Instructions 4-5: split stack pointer in two (stack pointer for callee done).
  % Instructions 6-9: Load the seal designated for this call.
  % Instruction 10: Seal capability for private stack to create data return pointer.
  % Instructions 11-13: Seal capability for return point to create code return pointer.
  % Instructions 14-15: Clean up temporary register and jump.
  % Instructions 16-26: The return code.
  % Instructions 16-23: check the stack base.
  % Instruction 24: join the stack pointer.
  % Instructions 25-26: pop 42 from the stack.
  The instructions for a $\scall{\offpc,\offsigma}{r_1}{r_2}$.
  The variable $\offpc$ is the offset from the first instruction of the call to the set of seals it uses.
  The variable $\offsigma$ is the offset in the set of seals to the call seal.
  The constant $\stkb$ is the globally agreed on stack base.
  There are a number of ``magic'' numbers in the code: line 1: $42$, garbage data to ensure a non-empty stack.
  Line 7: $-5$, offset to the first instruction of the call (from the point the capability was moved out of the $\pcreg$-register).
  Line 12: $5$, offset to the return address.
  Line 19: $5$, offset to fail.
  Line 21: offset to address after fail.
}
  \label{fig:call-code}
\end{figure}
